Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-10-26 cvx-flats-neurips21/cvx-flats.tex
Moving Lemma 1 up and cleaning other things: flats

\paragraph{The ``four quadrants'' of problem types.}
Above, we discuss a significant divergence in previous frameworks: constructing a surrogate given a \emph{target loss} versus a \emph{target statistic}.
In addition to the two possible targets, we may have one of two domains: a \emph{discrete} (i.e.\ finite) target prediction space, like a classification problem, or a \emph{continuous} one, like a regression or estimation problem.
We informally refer to the four resulting cases---target loss vs.\ target statistic, and discrete vs.\ continuous predictions---as the ``four quadrants'' of supervised learning problems, shown in Table~\ref{tab:quadrants}.
% For further examples, see Appendix~\ref{app:omitted-examples}.
In the context of these quadrants, Figure~\ref{fig:results-flow} gives a roadmap of our main results.

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{prop:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results for Q1, while simultaneously applying to other quadrants. We illustrate the breadth and power of $d$-flats by resolving two open questions for Q3 and Q4 in \S~\ref{sec:contin-consis}.}
\label{fig:results-flow}
\end{figure}
we introduce a notion of consistency in line with classical statistics \citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we define $L$ to be consistent if minimizing $L$ and applying a link function yields predictions converging to the correct statistic value.
%\jessiet{Do we still want this following paragraph?}
%\raft{Good question.  I edited a bit, but still up for discussion.}
%A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
The key observation which underpins our approach is that consistency for target losses is a special case of consistency for target statistics (\S~\ref{sec:consis-implies-indir}).
Therefore, property elicitation---which studies the exact minimizers of loss functions---allows us to give general lower bounds on prediction dimension of any convex surrogates corresponding to a target task; these bounds apply across all four quadrants.
See \S~\ref{subsec:complexity} for other prior work on notions of prediction dimension.

@@

In supervised learning, data is drawn from a distribution $\D$ over the space $\X \times \Y$ and the goal is to produce a hypothesis $f: \X \to \R$.
Here $\X$ is the \emph{feature space}, $\Y$ the \emph{label space}, and $\R$ the \emph{report} or \emph{prediction space}, possibly different from $\Y$.
%consider supervised learning problems in the space $\X \times \Y$, for some \emph{feature space} $\X$ and a \emph{label space} $\Y$, with data drawn from a distribution $D$ over $\X \times \Y$.
%The task is to produce a hypothesis $f: \X \to \R$, for some \emph{prediction space} $\R$, which may be different from $\Y$.
For example, in ranking problems, $\R$ may be all $|\Y|!$ permutations over the $|\Y|$ labels forming $\Y$.
We focus on surrogate losses, target problems, and their relationships to conditional distributions $p := \D_x = \Pr[Y|X=x]$ over $\Y$ given some $x \in X$.
We can often abstract away $x$, working directly with a set of (conditional) distributions over outcomes $\P \subseteq \simplex$, where $\simplex := \{p \in \reals^{|\Y|}_+ \mid \|p\|= 1\}$ is the probability simplex over labels.
We then write e.g.\ $\E_p \ell(r,Y)$ to mean the expected loss of prediction $r \in \R$ when $Y\sim p$.

@@

\raf{Lemma~\ref{lem:consistent-loss-implies-prop} goes here}

@@

\begin{theorem}\label{thm:mode}
  The mode has $\conscvx(\mode) = \eliccvx(\mode) = \infty$ (countably infinite) with respect to $\P$, the class of probability measures on $\Y=\reals$ with a smooth and bounded density and such that $\mode$ is single-valued.
\end{theorem}
\begin{proof}
  The proof of \citet[Theorem 1]{dearborn2020indirect} gives a distribution $p\in\P$ with $\mode(p) = 0 =: u$.
  It then introduces an arbitrary identification function $V:\hat\R\times\Y\to\reals^k$, $k\in\mathbb{N}$, and value $r\in\hat\R$ such that $p \in \zeros{V(r,\cdot)}$.
  Letting $F = \zeros{V(r,\cdot)}$, we therefore have an arbitrary $k$-flat containing $p$.
  The proof then proceeds to construct some $p'\in F$ with $\mode(p') \neq u$.
  Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} now gives $\conscvx(\mode) \geq \eliccvx(\mode) \geq k+1$.
  \btw{NOTE: the set $\P$ is not convex.  Take two univariance Gaussians centered at 0 and 1: they are each in $\P$ but their mixture is not.}
  As $k$ was arbitrary, the result follows.
\end{proof}

@@

\begin{ack}
  We thank Adam Bloniarz and Nishant Mehta for several key suggestions and insights.
  This material is based upon work supported by the National Science Foundation under Grant Nos.\ IIS-2045347 and 2040434. \jessie{confirm GRFP number}\raf{I think it's usually stated differently for GRFPs, since it's not really a grant application, right?}
\end{ack}

@@

Given a target loss $\ell$, we can define a statistic $\gamma$, the property it elicits.
Intuitively, consistency of a surrogate $L$ with respect to $\ell$ and $\gamma$ are equivalent, i.e. in both cases estimates should converge to values that minimize $\ell$-loss.
We formalize this by letting $\propdis$ be the $\ell$-regret, yielding Lemma~\ref{lem:consistent-loss-implies-prop}.
%If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent, as stated in Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.% for the appropriate choice of $\propdis$.

\raf{Important: I think we need to move this lemma up to \S~\ref{sec:informal-consis-implies-indir}.  Probably using this exact same statement.  Related: do we define $R_\ell$ anywhere?  Not in this section, unless I missed it...}
\begin{restatable}{lemma}{consistentlossimpliesprop}\label{lem:consistent-loss-implies-prop}
	Let a convex $\P \subseteq \simplex$ be given.
	%\jessiet{Defer to appendix?  Thoughts?}\raft{Let's keep it but move the proof.}
	Given a surrogate loss $L \in \L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
	$\mu(r,p) := R_\ell(r,p)$.
	Then there is a $\D$ such that
	$(L, \psi)$ is $\D$-consistent with respect to $\ell$ if and only if $(L,\psi)$ is $(\propdis, \D)$-consistent with respect to $\gamma := \prop{\ell}$. 
	\btw{Note: don't actually need $\gamma$ nonempty here.}
\end{restatable}

\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
	Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
	\begin{align}
	&\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf\nolimits_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	&\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	&\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
	Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
	Observe that the assumptions on $\L$ allow us to apply the Fubini-Tonelli Theorem~\cite[Theorem 2.37]{folland1999real}, which yields the equivalence of eq.~\ref{eqn:cons-loss-cond} to the next line.
\end{proof}

>>>>

\paragraph{The ``four quadrants'' of problem types.}
Above, we discuss a significant divergence in previous frameworks: constructing a surrogate given a \emph{target loss} versus a \emph{target statistic}.
In addition to the two possible targets, we may have one of two domains: a \emph{discrete} (i.e.\ finite) target prediction space, like a classification problem, or a \emph{continuous} one, like a regression or point estimation problem.
We informally refer to the four resulting cases---target loss vs.\ target statistic, and discrete vs.\ continuous predictions---as the ``four quadrants'' of supervised learning problems, shown in Table~\ref{tab:quadrants}.
% For further examples, see Appendix~\ref{app:omitted-examples}.
In the context of these quadrants, Figure~\ref{fig:results-flow} gives a roadmap of our main results.

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{prop:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results for Q1, while simultaneously applying to other quadrants. We illustrate the breadth and power of $d$-flats by resolving two open questions for Q3 and Q4 in \S~\ref{sec:contin-consis}.}
\label{fig:results-flow}
\end{figure}
we introduce a notion of consistency in line with classical statistics \citep{gyorfi2006distribution, fan1998efficient,ruppert1997local}.
Here we define $L$ to be consistent if minimizing $L$ and applying a link function yields predictions converging to the correct statistic value.
%\jessiet{Do we still want this following paragraph?}
%\raft{Good question.  I edited a bit, but still up for discussion.}
%A priori, it is not clear that compatible definitions of consistency could be given for both target statistics and target losses.
The key observation which underpins our approach is that consistency for target losses is a special case of consistency for target statistics (Lemma~\ref{lem:consistent-loss-implies-prop}). 
%(\S~\ref{sec:consis-implies-indir}).
Therefore, property elicitation---which studies the exact minimizers of loss functions---allows us to give general lower bounds on prediction dimension of any convex surrogates corresponding to a target task; these bounds apply across all four quadrants.
See \S~\ref{subsec:complexity} for other prior work on notions of prediction dimension.

@@

In supervised learning, data is drawn from a distribution $D$ over the space $\X \times \Y$ and the goal is to produce a hypothesis $f: \X \to \R$.
Here $\X$ is the \emph{feature space}, $\Y$ the \emph{label space}, and $\R$ the \emph{report} or \emph{prediction space}, possibly different from $\Y$.
%consider supervised learning problems in the space $\X \times \Y$, for some \emph{feature space} $\X$ and a \emph{label space} $\Y$, with data drawn from a distribution $D$ over $\X \times \Y$.
%The task is to produce a hypothesis $f: \X \to \R$, for some \emph{prediction space} $\R$, which may be different from $\Y$.
For example, in ranking problems, $\R$ may be all $|\Y|!$ permutations over the $|\Y|$ labels forming $\Y$.
We focus on surrogate losses, target problems, and their relationships to conditional distributions $p := D_x = \Pr[Y|X=x]$ over $\Y$ given some $x \in X$.
We can often abstract away $x$, working directly with a set of (conditional) distributions over outcomes $\P \subseteq \simplex$, where $\simplex := \{p \in \reals^{|\Y|}_+ \mid \|p\|= 1\}$ is the probability simplex over labels.
We then write e.g.\ $\E_p \ell(r,Y)$ to mean the expected loss of prediction $r \in \R$ when $Y\sim p$.

@@

We can now state Lemma~\ref{lem:consistent-loss-implies-prop}, with proof deferred to \S~\ref{sec:consis-implies-indir}.

\begin{restatable}{lemma}{consistentlossimpliesprop}\label{lem:consistent-loss-implies-prop}
	Let a convex $\P \subseteq \simplex$ be given.
	Given a surrogate loss $L \in \L$, link $\psi$, and target loss $\ell$, set
	$\mu(r,p) := \E_p \ell(r,Y) - \min_{r'}\E_p \ell(r',Y)$.
	Then there is a $\D$ such that
	$(L, \psi)$ is $\D$-consistent with respect to $\ell$ if and only if $(L,\psi)$ is $(\propdis, \D)$-consistent with respect to $\gamma := \prop{\ell}$. 
%	\btw{Note: don't actually need $\gamma$ nonempty here.}
\end{restatable}
\jessiet{Note: this is the formal version- do we want an informal version here instead, or to just leave the details in here?}

@@

\begin{theorem}\label{thm:mode}
  The mode has $\conscvx(\mode) = \eliccvx(\mode) = \infty$ (countably infinite) with respect to $\P$, the class of probability measures on $\Y=\reals$ with a smooth and bounded density and such that $\mode$ is single-valued.
\end{theorem}
\begin{proof}
  The proof of \citet[Theorem 1]{dearborn2020indirect} gives a distribution $p\in\P$ with $\mode(p) = 0 =: u$.
  It then introduces an arbitrary identification function $V:\hat\R\times\Y\to\reals^k$, $k\in\mathbb{N}$, and value $r\in\hat\R$ such that $p \in \zeros{V(r,\cdot)}$.
  Letting $F = \zeros{V(r,\cdot)}$, we therefore have an arbitrary $k$-flat containing $p$.
  The proof then proceeds to construct some $p'\in F$ with $\mode(p') \neq u$.
  Corollary~\ref{cor:Pcodim-flat-elic-relint-prop} now gives $\conscvx(\mode) \geq \eliccvx(\mode) \geq k+1$.
  \btw{NOTE: the set $\P$ is not convex.  Take two univariance Gaussians centered at 0 and 1: they are each in $\P$ but their mixture is not. \jessie{Should we be concerned about this}}
  As $k$ was arbitrary, the result follows.
\end{proof}

@@

\begin{ack}
  We thank Adam Bloniarz and Nishant Mehta for several key suggestions and insights.
  This material is based upon work supported by the National Science Foundation under Grant No.\ IIS-2045347. 
  Additionally, this material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. 2040434. \jessie{confirm GRFP number}
\end{ack}

@@

Given a target loss $\ell$, we can define a statistic $\gamma$, the property it elicits.
Intuitively, consistency of a surrogate $L$ with respect to $\ell$ and $\gamma$ are equivalent, i.e. in both cases estimates should converge to values that minimize $\ell$-loss.
We formalize this by letting $\propdis$ be the $\ell$-regret, $R_\ell := \E_p \ell(r,Y) - \min_{r'}\E_p \ell(r',Y)$, yielding Lemma~\ref{lem:consistent-loss-implies-prop}.
%If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent, as stated in Lemma~\ref{lem:consistent-loss-implies-prop}, proven in Appendix~\ref{app:misc-omitted-proofs}.% for the appropriate choice of $\propdis$.

\raf{Important: I think we need to move this lemma up to \S~\ref{sec:informal-consis-implies-indir}.  Probably using this exact same statement.  Related: do we define $R_\ell$ anywhere?  Not in this section, unless I missed it...}
\jessie{Good catch; I think we removed it from COLT to NeurIPS or vice versa}

\consistentlossimpliesprop*
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
	Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
	\begin{align}
	&\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf\nolimits_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	&\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	&\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
	Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
	Observe that the assumptions on $\L$ allow us to apply the Fubini-Tonelli Theorem~\cite[Theorem 2.37]{folland1999real}, which yields the equivalence of eq.~\ref{eqn:cons-loss-cond} to the next line.
\end{proof}
